---
title: "MAS Calling Script"
author: "Zachary Winn"
date: "2024-04-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup_markdown, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Requirments

This section checks if dependencies are installed, installs them if they are not present, and then library the packages required. 

```{r library_packages, echo=FALSE}
# Define a list of required packages
required_packages <- c("parallel", "foreach", "doParallel", "data.table", "R.utils", "knitr")

# Check if required packages are installed
missing_packages <- setdiff(required_packages, rownames(installed.packages()))

# Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load required packages
library(parallel)
library(foreach)
suppressMessages(library(doParallel))
suppressMessages(library(data.table))
library(knitr)
```

# Set Input Files

Here I am setting the names of my inputs.

```{r inputs, echo=FALSE}
# VCF file name
vcf_file <- "Example.vcf.gz"

# MAS key
mas_key_file <- "Example_File.txt"
```

# Read in Data and Convert to HapMap

Here I am using base R to read in a VCF and format it into a HapMap format to look at the allelic state of a the VCF markers.

```{r hapmap_conversion, echo=FALSE}
# Count lines that start with ##
count_header_lines <- function(x) {
  # Open connection to file
  con <- file(x, "r")
  # Set the number for first line
  header_lines <- 0
  # While 
  while (TRUE) {
    # Read in the line = 1
    line <- readLines(con, n = 1)
    # If the line equal to zero 
    if (length(line) == 0) break  # Exit loop when end of file is reached
    # If the line contains ##
    if (substr(line, 1, 2) == "##") {
      # Add to length
      header_lines <- header_lines + 1
    } else {
      break  # Exit loop when non-header lines start
    }
  }
  close(con)
  return(header_lines)
}

# Read vcf as table
vcf <- as.data.frame(data.table::fread(input = vcf_file, skip = count_header_lines(vcf_file)))

# Remove the first row if it was a header row
vcf[,10:ncol(vcf)] <- lapply(vcf[,10:ncol(vcf)], function(x){substr(x, 1, 3)}) 

# Read in MAS key 
key <- read.table(mas_key_file, sep = "\t", header = TRUE, na.strings = c("", "NA"))

# Key markers
key_markers <- key[,"Markers"]

# Sting separate and unique markers
key_markers <- unique(unlist(strsplit(key_markers, split = ":")))

# subset
vcf <- vcf[vcf[,"ID"] %in% key_markers,]

# Define a function to be executed in parallel
vcf_to_hapmap <- function(x) {
  
  temp1 <- x
  temp2 <- temp1[,"REF"]
  temp3 <- temp1[,"ALT"]
  temp4 <- temp1[,"ID"]
  temp5 <- temp1
  temp1 <- temp1[,10:ncol(temp1)]
  
  if(temp2 %in% c("A", "T", "G", "C") & temp3 %in% c("A", "T", "G", "C")){

    #print("SNP")
    temp1[,1:ncol(temp1)] <- lapply(temp1[,1:ncol(temp1)], function(x){gsub(0, temp2, x)})
    temp1[,1:ncol(temp1)] <- lapply(temp1[,1:ncol(temp1)], function(x){gsub(1, temp3, x)})
    temp1[temp1=="A/A"] = "A"
    temp1[temp1=="T/T"] = "T"
    temp1[temp1=="C/C"] = "C"
    temp1[temp1=="G/G"] = "G"
    temp1[temp1=="A/G"] = "R"
    temp1[temp1=="G/A"] = "R"
    temp1[temp1=="C/T"] = "Y"
    temp1[temp1=="T/C"] = "Y"
    temp1[temp1=="G/C"] = "S"
    temp1[temp1=="C/G"] = "S"
    temp1[temp1=="A/T"] = "W"
    temp1[temp1=="T/A"] = "W"
    temp1[temp1=="G/T"] = "K"
    temp1[temp1=="T/G"] = "K"
    temp1[temp1=="A/C"] = "M"
    temp1[temp1=="C/A"] = "M"
    temp1[temp1=="./."] = "N"
    temp1<-data.frame(`rs#` = temp4,
                      alleles = paste(temp2,"/",temp3,sep = ""),
                      chrom = temp5[,"#CHROM"],
                      pos = temp5[,"POS"],
                      strand = "+",
                      `assembly#` = "NA",
                      center = "NA",
                      protLSID = "NA",
                      assayLSID = "NA",
                      panelLSID = "NA",
                      QCcode = "NA",
                      temp1,
                      check.names = FALSE,
                      row.names = NULL)
    return(temp1)
    remove(temp1, temp2, temp3, temp4, temp5)
    
  }else if (nchar(temp2)<nchar(temp3) | nchar(temp2)>nchar(temp3)){
        
    #print("SNP")
    temp1[,1:ncol(temp1)] <- lapply(temp1[,1:ncol(temp1)], function(x){gsub(0, temp2, x)})
    temp1[,1:ncol(temp1)] <- lapply(temp1[,1:ncol(temp1)], function(x){gsub(1, temp3, x)})
    
    if(nchar(temp2)>nchar(temp3)){

      # string
      string <- paste(temp2, temp2, sep = "/")
      
      # replace
      temp1[temp1==string] <- substr(temp2, start = nchar(temp3)+1, stop = nchar(temp3)+1)
      
      # string
      string <- paste(temp2, temp3, sep = "/")
      
      # replace
      temp1[temp1==string] <- 0
      
      # string
      string <- paste(temp3, temp2, sep = "/")
      
      # replace
      temp1[temp1==string] <- 0
      
      # string
      string <- paste(temp3, temp3, sep = "/")
      
      # replace
      temp1[temp1==string] <- "-"
      
      # replace
      temp1[temp1=="./."] <- "N"
      
      # temp1
      temp1<-data.frame(`rs#` = temp4,
                        alleles = paste(substr(temp2, start = nchar(temp3)+1, stop = nchar(temp3)+1),"/-",sep = ""),
                        chrom = temp5[,"#CHROM"],
                        pos = temp5[,"POS"],
                        strand = "+",
                        `assembly#` = "NA",
                        center = "NA",
                        protLSID = "NA",
                        assayLSID = "NA",
                        panelLSID = "NA",
                        QCcode = "NA",
                        temp1,
                        check.names = FALSE,
                        row.names = NULL)
    
    return(temp1)
    remove(temp1, temp2, temp3, temp4, temp5, string)         
        
    }else if(nchar(temp2)<nchar(temp3)){
      
      # string
      string <- paste(temp3, temp3, sep = "/")
      
      # replace
      temp1[temp1==string] <- substr(temp3, start = nchar(temp2)+1, stop = nchar(temp2)+1)
      
      # string
      string <- paste(temp2, temp3, sep = "/")
      
      # replace
      temp1[temp1==string] <- 0
      
      # string
      string <- paste(temp3, temp2, sep = "/")
      
      # replace
      temp1[temp1==string] <- 0
      
      # string
      string <- paste(temp2, temp2, sep = "/")
      
      # replace
      temp1[temp1==string] <- "-"
      
      # replace
      temp1[temp1=="./."] <- "N"
      
      # return          
      temp1<-data.frame(`rs#` = temp4,
                        alleles = paste("-/",substr(temp3, start = nchar(temp2)+1, stop = nchar(temp2)+1),sep = ""),
                        chrom = temp5[,"#CHROM"],
                        pos = temp5[,"POS"],
                        strand = "+",
                        `assembly#` = "NA",
                        center = "NA",
                        protLSID = "NA",
                        assayLSID = "NA",
                        panelLSID = "NA",
                        QCcode = "NA",
                        temp1,
                        check.names = FALSE,
                        row.names = NULL)
      
      return(temp1)
      remove(temp1, temp2, temp3, temp4, temp5)   

    }
  }
}

# Set the number of cores to use
num_cores <- detectCores()-1

# Create a cluster
cl <- makeCluster(num_cores)

# Register the cluster
registerDoParallel(cl)

# Message
print_text <- paste("### HapMap conversion start:", Sys.time(), "###") 

# Print
print(paste(rep("#", nchar(print_text)), collapse = ""))
print(print_text)
print(paste(rep("#", nchar(print_text)), collapse = ""))

# For each
hapmap <- foreach(i = 1:nrow(vcf), .combine = 'rbind') %dopar% {
  # Pull data
  temp1 <- vcf[i,]
  
  # Do function
  temp1 <- vcf_to_hapmap(temp1)
  
  # Return 
  return(temp1)
  
  # Remove
  remove(temp1)
}

# Stop the cluster
stopCluster(cl)

# Remove cluster
remove(cl)

# Write out hapmap
fwrite(as.data.table(hapmap, keep.rownames = FALSE), 
       file = paste(strsplit(vcf_file, "\\.")[[1]][1],"_hapmap.hmp.txt.gz", sep = ""),
       sep = "\t")

# Message
print_text <- paste("### HapMap conversion stop:", Sys.time(), "###") 

# Print
print(paste(rep("#", nchar(print_text)), collapse = ""))
print(print_text)
print(paste(rep("#", nchar(print_text)), collapse = ""))
```

# Get Report

Here I am taking the haplotyping file and writing out reports.

```{r making_reports, echo=FALSE}
# make object
auto_suggestions <- c()

# For loop
for (i in unique(key[,"Locus"])){
  
  # Pull information
  temp1 <- key[key[,"Locus"]==i,]
  
  # Get markers
  temp2 <-  unlist(strsplit(temp1[,"Markers"], ":"))
  
  # Pull markers
  temp3 <- as.data.frame(hapmap[hapmap[,"rs#"] %in% temp2, ])
  rownames(temp3) <- temp3[,1]
  temp3 <- temp3[,12:ncol(temp3)]
  
  # Check if markers are found in both files
  if(all(temp2 %in% rownames(temp3))==FALSE){
    message("#################################")
    message("Markers identified for ",
            i,
            " are not found HapMap.",
            " Not producing call for locus ",
            i,
            " and moving to the next locus.",
            " If this is in error, check your VCF input marker names and your haplotyping key.")
    next
  }
  
  # Pull haplotypes
  if(is.na(temp1[,"Pos"])){temp4 <- NA}else{temp4 <- unlist(strsplit(temp1[,"Pos"], "_"))}
  if(is.na(temp1[,"Het"])){temp5 <- NA}else{temp5 <- unlist(strsplit(temp1[,"Het"], "_"))}
  if(is.na(temp1[,"Neg"])){temp6 <- NA}else{temp6 <- unlist(strsplit(temp1[,"Neg"], "_"))}
  if(is.na(temp1[,"Missing"])){temp7 <- NA}else{temp7 <- unlist(strsplit(temp1[,"Missing"], "_"))}
  
  # Check if auto call is requested
  if(temp1[,"Auto"]==TRUE){
    
    # Check if positive calls are given
    if(is.na(temp4[1])){
      stop("It appears there are no positive haplotypes given for ", 
           i,
           ". There must be at least one positive haplotype defined for the 'Auto' feature.",
           " Please check your inputs and resubmit!")
    }
    
    # Pull number of missing
    n.missing <- temp1[,"N_Missing"]
    
    # Check if number of missing is defined
    if(is.na(n.missing) | !is.numeric(n.missing)){
      stop("Number of tolerated missing calls must be defined and must be a number for ",
           i,
           "! Check the Column 'N_Missing' and resubmit!")
    }
    
    # Check if the locus only has one marker 
    if(length(temp2)==1){
      stop("Only one defining marker detected for ",
           i,
           ". The 'Auto' function is only available for loci with multiple markers! ",
           "Please define your haplotypes, set auto to 'FALSE', and resubmit.")
    }
     
    # Get ref alt of each marker in iupac
    ref_alt <- c()
    
    # Permutate through markers
    for(j in rownames(temp3)){
      temp1.1 <- as.character(temp3[j,])
      temp1.1 <- unique(temp1.1)
      temp1.2 <- temp1.1[grep("G",temp1.1)]
      temp1.3 <- temp1.1[grep("T",temp1.1)]
      temp1.4 <- temp1.1[grep("C",temp1.1)]
      temp1.5 <- temp1.1[grep("A",temp1.1)]
      temp1.6 <- temp1.1[grep("-",temp1.1)]
      temp1.1 <- data.frame(Marker=j,
                            G=ifelse(is.null(temp1.2), NA, temp1.2),
                            `T`=ifelse(is.null(temp1.3), NA, temp1.3),
                            C=ifelse(is.null(temp1.4), NA, temp1.4),
                            A=ifelse(is.null(temp1.5), NA, temp1.5),
                            gap=ifelse(is.null(temp1.6), NA, temp1.6))
      temp1.1 <- temp1.1[, colSums(is.na(temp1.1)) != nrow(temp1.1)]
      # Check if the maker is biallelic
      if(ncol(temp1.1[, colSums(is.na(temp1.1)) != nrow(temp1.1)])!=3 & ncol(temp1.1[, colSums(is.na(temp1.1)) != nrow(temp1.1)])!=2){
        stop("Marker ", 
             j, 
             " associated with locus ", 
             i, 
             " appears to be multiallelic rather than biallelic.",
             " Calls can only be made in Auto mode with biallelic markers!",
             " Please check your inputs and resubmit!")
      }else if(ncol(temp1.1[, colSums(is.na(temp1.1)) != nrow(temp1.1)])==2){
        message("#################################")
        message("Marker ", 
                j, 
                " associated with locus ", 
                i, 
                " appears to be monomorphic rather than biallelic.",
                " Moving forward with assumption that the marker is dominant...")
        temp1.1 <- cbind(temp1.1, NA)
        colnames(temp1.1) <- c("Marker", "Allele_1", "Allele_2")
        ref_alt <- rbind(ref_alt, temp1.1)
        remove(temp1.1, temp1.2, temp1.3, temp1.4, temp1.5, temp1.6)
      }else{
        colnames(temp1.1) <- c("Marker", "Allele_1", "Allele_2")
        ref_alt <- rbind(ref_alt, temp1.1)
        remove(temp1.1, temp1.2, temp1.3, temp1.4, temp1.5, temp1.6) 
      }
    }
  
    # assign het call
    ref_alt$Allele_Het <- paste(ref_alt[,"Allele_1"],
                                "/",
                                ref_alt[,"Allele_2"],
                                sep = "")
    ref_alt[ref_alt=="A/G"] = "R"
    ref_alt[ref_alt=="G/A"] = "R"
    ref_alt[ref_alt=="C/T"] = "Y"
    ref_alt[ref_alt=="T/C"] = "Y"
    ref_alt[ref_alt=="G/C"] = "S"
    ref_alt[ref_alt=="C/G"] = "S"
    ref_alt[ref_alt=="A/T"] = "W"
    ref_alt[ref_alt=="T/A"] = "W"
    ref_alt[ref_alt=="G/T"] = "K"
    ref_alt[ref_alt=="T/G"] = "K"
    ref_alt[ref_alt=="A/C"] = "M"
    ref_alt[ref_alt=="C/A"] = "M"
    
    # assign dominant calls
    ref_alt[ref_alt=="A/NA"] = NA
    ref_alt[ref_alt=="G/NA"] = NA
    ref_alt[ref_alt=="C/NA"] = NA
    ref_alt[ref_alt=="T/NA"] = NA
    
    # Define positive allele
    positive <- data.frame(Haplotype = temp4)
    positive <- cbind(positive, matrix(data = unlist(strsplit(temp4, split = ":")),
                                       nrow = nrow(positive),
                                       byrow = TRUE))
    positive[positive == "N"] <- NA
    positive[positive == "R"] <- NA
    positive[positive == "Y"] <- NA
    positive[positive == "S"] <- NA
    positive[positive == "W"] <- NA
    positive[positive == "K"] <- NA
    positive[positive == "M"] <- NA
    positive <- na.omit(positive)
    colnames(positive) <- c("Haplotype", ref_alt[,"Marker"])
    
    if(nrow(positive)==0){
      stop("It appears that no positive cases for ",
           i,
           " are complete case!",
           " Positive cases must contain at least one complete, homozygous sequence (e.g., A:G:T).",
           " Check your inputs and resubmit!")
    }
    
    # Define negative allele 
    negative <- positive
    
    for(j in colnames(positive)[2:ncol(positive)]){
      
      # pull allele 
      temp1.1 <- negative[,j]
      
      # pull corresponding marker
      temp1.2 <- ref_alt[ref_alt[,"Marker"]==j,]
      
      # identify alternate 
      temp1.3 <- ifelse(temp1.2[,"Allele_1"]==temp1.1, temp1.2[,"Allele_2"], temp1.2[,"Allele_1"])
      
      # replace NA
      temp1.3 <- ifelse(is.na(temp1.3), "N", temp1.3)
      
      # rename
      negative[,j] <- temp1.3
      
      # remove
      remove(temp1.1, temp1.2, temp1.3)
      
    }
  
    # Define heterozygus cases
    heterozygous <- as.character(ref_alt[,"Allele_Het"])
    heterozygous <- paste(heterozygous, collapse = ":")
    heterozygous <- data.frame(Haplotype=heterozygous,
                               t(ref_alt[,"Allele_Het"]))
    colnames(heterozygous)[2:ncol(heterozygous)] <- ref_alt[,"Marker"]
    
    # if there are NAs which indicate that there is a dominant marker
    if(length(unique(is.na(heterozygous)))!=1){
      heterozygous<-c()
      for(j in 1:nrow(ref_alt)){
        if(is.na(ref_alt[j,"Allele_Het"])){
          heterozygous<-c(heterozygous,ref_alt[j,"Allele_1"])
        }else{
          heterozygous<-c(heterozygous,ref_alt[j,"Allele_Het"])
        }
      }
      heterozygous<-data.frame(Haplotype=paste(heterozygous, collapse = ":"),
                               t(heterozygous))
      colnames(heterozygous)[2:ncol(heterozygous)] <- ref_alt[,"Marker"]
    }
    
    # Define missing case
    missing <- data.frame(Haplotype = paste(rep("N", length(2:ncol(heterozygous))), collapse = ":"),
                          t(rep("N", length(2:ncol(heterozygous)))))
    colnames(missing)[2:ncol(missing)] <- ref_alt[,"Marker"]
    
    # check if negative cases have N
    if("N" %in% negative[,2:ncol(negative)]){
      temp1.1 <- negative[,2:ncol(negative)]
      temp1.1 <- colnames(temp1.1[,temp1.1=="N"])
      dominant_markers <- temp1.1
      remove(temp1.1)
    }
    
    # Combine the cases
    cases <- data.frame(Type=c("Positive", "Heterozygous", "Negative", "Missing"),
                        rbind(positive, heterozygous, negative, missing))
    
    # Get every other possible combination
    haplotypes <- expand.grid(cases[,3:ncol(cases)])
    haplotypes[,1:ncol(haplotypes)] <- lapply(haplotypes[,1:ncol(haplotypes)], as.character) 
    haplotypes <- unique(haplotypes)
    
    # make a list of deinfed haplotypes
    for(j in 1:nrow(haplotypes)){
      temp1.1 <- haplotypes[j,]
      temp1.1 <- paste(temp1.1[,1:length(temp1.1)], collapse = ":")
      if(j==1){
        temp1.2<-temp1.1
      }else{
        temp1.2<-c(temp1.2, temp1.1)
      }
      remove(temp1.1)
    }
    
    haplotypes$Haplotype<-temp1.2
    remove(temp1.2)
    
    # Give an innitial call to things which are defined
    haplotypes[,"Call"] <- ifelse(haplotypes[,"Haplotype"] %in% temp4, "POS",
                              ifelse(haplotypes[,"Haplotype"] %in% temp5, "HET",
                                     ifelse(haplotypes[,"Haplotype"] %in% temp6, "NEG",
                                            ifelse(haplotypes[,"Haplotype"] %in% temp7, "MISSING", NA))))
    
    #### If the sequence contains X number of N's call it missing
    
    # Empty vector
    N_Missing <- c()
    
    # Count number of N in a string
    for (j in 1:nrow(haplotypes)) {
      N_Missing <- c(N_Missing, sum(unlist(gregexpr("N", haplotypes[j, "Haplotype"], fixed = TRUE)) > 0))
    }
    
    # Report that 
    haplotypes[,"N_Missing"] <- N_Missing
    
    # Assign missing if N is greater than n.missing
    haplotypes[,"Call"] <- ifelse(haplotypes[,"N_Missing"]>n.missing, "MISSING", haplotypes[,"Call"])
    
    ### If the sequence is the same as the positive case but with a het call it het
    
    # make a list for heterozygous cases
    check<-expand.grid(cases[c(1,2,4),c(3:ncol(cases))])
    
    # Make sure they are in correct format
    check[,1:ncol(check)]<-lapply(check[,1:ncol(check)], as.character)
    
    # Make haplotypes
    for(j in 1:nrow(check)){
      temp1.1 <- check[j,]
      temp1.1 <- paste(temp1.1[,1:length(temp1.1)], collapse = ":")
      if(j==1){
        temp1.2<-temp1.1
      }else{
        temp1.2<-c(temp1.2, temp1.1)
      }
      remove(temp1.1)
    }
    
    # Bind in haplotypes
    check[,"Haplotype"] <- temp1.2
  
    # Remove
    remove(temp1.2)
    
    # Remove any cases that are already defined
    temp1.1 <- haplotypes[!is.na(haplotypes[,"Call"]),"Haplotype"]
    check <- check[!check[,"Haplotype"] %in% temp1.1, ]
    remove(temp1.1)
    
    # Redefine
    haplotypes[haplotypes[,"Haplotype"] %in% check[,"Haplotype"],"Call"] <- "HET"
    
    #### If the sequence contains at least one alternate of the positive case call it negative
    
    # make a list for negative cases
    check<-expand.grid(cases[c(1,2,3,4),c(3:ncol(cases))])
    
    # Make sure they are in correct format
    check[,1:ncol(check)]<-lapply(check[,1:ncol(check)], as.character)
    
    # Make haplotypes
    for(j in 1:nrow(check)){
      temp1.1 <- check[j,]
      temp1.1 <- paste(temp1.1[,1:length(temp1.1)], collapse = ":")
      if(j==1){
        temp1.2<-temp1.1
      }else{
        temp1.2<-c(temp1.2, temp1.1)
      }
      remove(temp1.1)
    }
    
    # Bind in haplotypes
    check[,"Haplotype"] <- temp1.2
    
    # Remove
    remove(temp1.2)
    
    # Remove any cases that are already defined
    temp1.1 <- haplotypes[!is.na(haplotypes[,"Call"]),"Haplotype"]
    check <- check[!check[,"Haplotype"] %in% temp1.1, ]
    remove(temp1.1)
    
    # Redefine
    if(exists("dominant_markers")){
      check[,dominant_markers]=NA 
      temp1.1 <- na.omit(check)
      temp1.2 <- check[!complete.cases(check),]
      haplotypes[haplotypes[,"Haplotype"] %in% temp1.1[,"Haplotype"],"Call"] <- "NEG"
      haplotypes[haplotypes[,"Haplotype"] %in% temp1.2[,"Haplotype"],"Call"] <- "MISSING"
      remove(temp1.1, temp1.2, dominant_markers)
    }else{
      haplotypes[haplotypes[,"Haplotype"] %in% check[,"Haplotype"],"Call"] <- "NEG"
    }
    
    # Reconcatinate and report
    for (j in unique(haplotypes[,"Call"])){
      temp1.1 <- haplotypes[haplotypes[,"Call"]==j,]
      temp1.1 <- as.character(temp1.1[,"Haplotype"])
      temp1.1 <- paste(temp1.1, collapse = "_")
      if(j=="POS"){
        temp1[,"Pos"]<-temp1.1
        remove(temp1.1)
      }else if(j=="HET"){
        temp1[,"Het"]<-temp1.1
        remove(temp1.1)
      }else if(j=="NEG"){
        temp1[,"Neg"]<-temp1.1
        remove(temp1.1)
      }else if(j=="MISSING"){
        temp1[,"Missing"]<-temp1.1
        remove(temp1.1)
      }
    }
    
    # Pull haplotypes again
    if(is.na(temp1[,"Pos"])){temp4 <- NA}else{temp4 <- unlist(strsplit(temp1[,"Pos"], "_"))}
    if(is.na(temp1[,"Het"])){temp5 <- NA}else{temp5 <- unlist(strsplit(temp1[,"Het"], "_"))}
    if(is.na(temp1[,"Neg"])){temp6 <- NA}else{temp6 <- unlist(strsplit(temp1[,"Neg"], "_"))}
    if(is.na(temp1[,"Missing"])){temp7 <- NA}else{temp7 <- unlist(strsplit(temp1[,"Missing"], "_"))}
    
    # add to auto suggestions
    auto_suggestions<-rbind(auto_suggestions, temp1)
    
    # Remove
    remove(cases, check, haplotypes, heterozygous, missing, negative, positive, ref_alt)
    
  }
  
  # For every individual in the hapmap
  for (j in colnames(temp3)){
    
    # Pull markers for individual
    temp8 <- temp3[,j]
    
    # Pull markers and collaps into a haplotype
    temp9 <- paste(temp8, collapse = ":")
      
    # Make call
    temp10 <- ifelse(temp9 %in% temp4, "POS",
                     ifelse(temp9 %in% temp5, "HET",
                            ifelse(temp9 %in% temp6, "NEG",
                                   ifelse(temp9 %in% temp7, NA, "UNIDENTIFIED"))))
    
        
    
    # Format into a dataframe
    temp11 <- data.frame(Line = j,
                         t(temp8),
                         temp9, 
                         temp10)
    
    # Pull optional annotations
    temp13 <- temp1[,"Pos_Annot"]
    temp14 <- temp1[,"Het_Annot"]    
    temp15 <- temp1[,"Neg_Annot"]
    temp16 <- temp1[,"Missing_Annot"]
    
    # Format annotations
    if(is.na(temp11[,ncol(temp11)])==TRUE & is.na(temp16)==FALSE){
      temp11[,ncol(temp11)] <- temp16
    }else if(temp11[,ncol(temp11)]=="POS" & is.na(temp13)==FALSE){
      temp11[,ncol(temp11)] <- temp13
    }else if(temp11[,ncol(temp11)]=="HET" & is.na(temp14)==FALSE){
      temp11[,ncol(temp11)] <- temp14
    }else if(temp11[,ncol(temp11)]=="NEG" & is.na(temp15)==FALSE){
      temp11[,ncol(temp11)] <- temp15
    }
    
    # Remove
    remove(temp13, temp14, temp15, temp16)
    
    # Make new column names
    colnames(temp11) <- c("Line", rownames(temp3), paste(i, "_Haplotype", sep = ""), paste(i, "_Summary", sep = ""))
    
    # Check if this is the first line in the hapmap
    if(j == colnames(temp3)[1]){
      
      # Make a new dataframe
      temp12<-temp11
      
      # Remove temporary files
      remove(temp8, temp9, temp10, temp11)
      
    }else{
      
      # Bind onto that new dataframe
      temp12<-rbind(temp12, temp11)
      
      # Remove temporary files
      remove(temp8, temp9, temp10, temp11)
      
    }
  }
  
  # Check if this is the first locus to be predicted
  if(i == unique(key[,"Locus"])[1]){
    
    # Make a new dataframe
    marker_report<-temp12
    
    # Remove temporary files
    remove(temp1, temp2, temp3, temp4, temp5, temp6, temp7, temp12)
    
  }else{
    
    # Else bind onto that new dataframe
    marker_report<-merge(marker_report, temp12, by="Line")
    
    # Remove temporary files
    remove(temp1, temp2, temp3, temp4, temp5, temp6, temp7, temp12)
  }
}

# Write out full report
write.csv(marker_report,
          paste(strsplit(vcf_file, "\\.")[[1]][1],"_Marker_Report_Full.csv",sep = ""),
          row.names = FALSE,
          quote = TRUE)

# Pull column names
columns_to_pull <- c("Line", colnames(marker_report)[grep("Summary", colnames(marker_report))])

# Pull summary and rename
marker_summary<-marker_report[,columns_to_pull]
colnames(marker_summary)<-gsub("_Summary", "", colnames(marker_summary))

# Write out just summary
write.csv(marker_summary,
          paste(strsplit(vcf_file, "\\.")[[1]][1],"_Marker_Report_Summary.csv",sep = ""),
          row.names = FALSE,
          quote = TRUE)

# Get information
missing_haplotypes <- c()

# Pull out the problem haplotypes
for (i in unique(key[,"Locus"])){
 
  # Pull data
  temp1 <- as.data.frame(marker_report[,c("Line",colnames(marker_report)[grep(pattern = i, colnames(marker_report))])])
  
  # if the locus isnt reported, skip over
  if(ncol(temp1)==1){
    next
  }
  
  # Rename columns
  temp2 <- colnames(temp1)
  temp2 <- gsub(paste(i, "_", sep = ""), "", temp2)
  colnames(temp1) <- temp2
  
  # Filter
  temp1 <- na.omit(temp1[temp1[,"Summary"]=="UNIDENTIFIED",])
  
  # If
  if(nrow(temp1)!=0){
  
    # Pull data
    temp3 <- key[key[,"Locus"]==i,]
    
    # Make dataframe
    temp4 <- unique(temp1[,"Haplotype"])
    
    # Bind
    temp5 <- data.frame(temp3, 
                        Unidentified_Haplotype=temp4,
                        row.names = NULL)
    
    # Rbind
    missing_haplotypes <- rbind(missing_haplotypes, temp5)
    
    # Remove
    remove(temp3, temp4, temp5)
      
  }
  
  # Remove
  remove(temp1, temp2)
  
}

# Check if a file needs to be written out
if(!is.null(missing_haplotypes)){
 
  # Write out information
  write.csv(missing_haplotypes,
            paste(strsplit(vcf_file, "\\.")[[1]][1],"_Unidentified_Haplotypes.csv",sep = ""),
            row.names = FALSE,
            quote = TRUE)  
  
}

# Check if a file needs to be written out
if(!is.null(auto_suggestions)){
 
  # Write out information
  write.csv(auto_suggestions,
            paste(strsplit(vcf_file, "\\.")[[1]][1],"_Auto_Suggested_Haplotypes.csv",sep = ""),
            row.names = FALSE,
            quote = TRUE)  
  
}
```